---
title: "Fitting Insurance Claims Severity Data with Coverage Modifications"
subtitle: How to Trick `fitdistrplus::fitdist` to Work with Customized Density Functions
author: | 
  | [Haoen Cui](https://haoen-cui.github.io/)
date: | 
  | May 30, 2020 
  | [satRday Chicago](https://chicago2020.satrdays.org/)
  | [Presentation GitHub Link](https://github.com/Haoen-Cui/talks/tree/master/claims-data-with-coverage-modification)
output:
  beamer_presentation:
    colortheme: "seahorse"
    slide_level: 2
    toc: true
header-includes:
  # short title for sidebar, headers, and footers 
  - \AtBeginDocument{\title[Coverage Modifications]{Fitting Insurance Claims Severity Data with Coverage Modifications}}
  # small font for code chunks 
  - \AtBeginDocument{\renewenvironment{Shaded} {\begin{snugshade}\footnotesize} {\end{snugshade}}}
linkcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Distribution Fitting 

## General Framework 

$$
\boxed{
\mathcal{P}(\theta)
    \underset{\text{
        Parameter Fitting
    }}{\overset{\text{
        Data Generating Process
    }}{\rightleftharpoons}}
\mathcal{D} = \{x_i\}_{i=1}^n
}
$$

- Data generating process 
    - Given distribution model $X \sim \mathcal{P}(\theta)$
    - Generate i.i.d. data $\mathcal{D} = \{x_i\}_{i=1}^n$ from $\mathcal{P}$

- Fitting parametric distribution model
    - Given observed data $\mathcal{D} = \{x_i\}_{i=1}^n$, assuming i.i.d.
    - Find parameter $\widehat{\theta}$ such that distribution model $\mathcal{P}(\widehat{\theta})$ "*fits best*"
        - Maximum Likelihood Estimation 
        - Generalized Method of Moments
        - Maximum Goodness-of-Fit Estimation

## Distribution Fitting Example 

Suppose we have an unknown data generating process ^[Contaminated normal distribution used for data simulation: $$ (1 - p) \cdot \mathcal{N}(\mu, {\sigma_1}^2) + p \cdot \mathcal{N}(\mu, {\sigma_2}^2) $$], and we attempt to find a set of parameters $(\mu^*, \sigma^*)$ 
such that 
$$\mathcal{N}(\mu^*, {\sigma^*}^2)$$ 
*fits* the data the best (among normal distribution family).

```{r dist_fit_data, echo=FALSE}
# parameters 
p <- 0.05 
mu <- 0 
sigma1 <- 1
sigma2 <- 3
n_sample <- 500 

# data simulation 
n_contaminated <- rbinom(n = 1, size = n_sample, prob = p) 
x <- c(
    rnorm(n = n_sample - n_contaminated, mean = mu, sd = sigma1), 
    rnorm(n = n_contaminated, mean = mu, sd = sigma2)
)
```

```{r dist_fit_data_summary}
summary(x)
```

## Distribution Fitting Example (Cont.)

```{r dist_fit_ex}
fit <- fitdistrplus::fitdist(x, distr = dnorm, method = 'mle')
plot(fit)
```

# Coverage Modifications 

## Incurred Raw Loss v.s. Insurer Covered Loss 

When an *insured* incurs a loss of amount $x$, part of the raw amount will be covered by the *insurer* (insurance policy). The process of computing covered loss from raw loss is known as *coverage modification* in actuarial mathematics. Some typical configurations are listed below:

- (Ordinary) Policy Deductible $d \geq 0$
$$
x \mapsto (x - d)_{+}
$$
- Policy Limit $u > d$
$$
x \mapsto \min\{x, u\}
$$
- Coinsurance Factor $\alpha \in [0, 1]$
$$
x \mapsto \alpha \cdot x
$$

## Coverage Modifications at a Glance

```{r cov_mod_helper, echo=FALSE}
modify_loss <- function(
    loss, 
    deductible = 0, 
    franchise = FALSE, 
    limit = Inf, 
    coinsurance = 1, 
    inflation = 0, 
    per.loss = FALSE
) {
  # 
  # a helper function to compute modified (covered) loss amount
  #   arguments align with actuar::coverage in version 2.3-1
  # 
  # Arguments: 
  #   - loss: a 1-D numeric vector for raw (unmodified) loss amounts
  #   - deductible: a unique positive numeric value; policy deductible
  #   - franchise: logical; 
  #                TRUE for a franchise deductible, 
  #                FALSE (default) for an ordinary deductible.
  #   - limit: a unique positive numeric value larger than deductible; policy limit
  #   - coinsurance: a unique value between 0 and 1; the proportion of coinsurance
  #   - inflation: a unique value between 0 and 1; the rate of inflation
  #   - per.loss: logical; 
  #               TRUE for the per loss distribution, 
  #               FALSE (default) for the per payment distribution
  # 
  # Value:
  #   a 1-D numeric vector containing covered loss amounts
  # 
  
  # inflation (and enforce type conversion)
  loss.mod <- (1.0 + inflation) * as.numeric(loss)
  # limit 
  loss.mod <- sapply(loss.mod, function(loss) min(loss, limit))
  # deductible 
  loss.mod <- 
    sapply(loss.mod - deductible, function(loss) max(loss, 0)) + 
    ifelse(franchise, deductible, 0)
  # left missing data type: truncation (per payment) or censoring (per loss)
  if (!per.loss) { loss.mod <- loss.mod[(loss.mod > 0)] }
  # coinsurance factor
  return(loss.mod * coinsurance)
}
```

```{r cov_mod_param, echo=FALSE}
DEDUCTIBLE <- 100 
LIMIT <- 250  
COINSURANCE <- 0.4
LOSS_MIN <- 0 
LOSS_MAX <- LIMIT + 50 
```

```{r cov_mod_plot_line, echo=FALSE, message=FALSE, warning=FALSE}
raw_loss <- seq(from = LOSS_MIN, to = LOSS_MAX, by = 1)
covered_loss <- modify_loss(
    loss = raw_loss, 
    deductible = DEDUCTIBLE, 
    limit = LIMIT, 
    coinsurance = COINSURANCE, 
    per.loss = TRUE 
)

p1 <- ggplot2::ggplot(
    data = data.frame(
        'x' = raw_loss, 
        'y' = covered_loss, 
        'not_payment' = raw_loss <= DEDUCTIBLE
    )
) + 
    ggplot2::geom_line(mapping = ggplot2::aes(x = x, y = y, linetype = not_payment)) + 
    ggplot2::geom_vline(xintercept = DEDUCTIBLE, linetype = 'dotted') + 
    ggplot2::geom_vline(xintercept = LIMIT, linetype = 'dotted') + 
    # ggplot2::coord_fixed() + 
    ggplot2::labs(x = 'Incurred Raw Loss', y = 'Insurer Covered Loss') + 
    ggplot2::theme_bw() + 
    # do not show legend 
    ggplot2::theme(legend.position = "none")

raw_loss <- rnorm(n = n_sample, mean = 150, sd = 30)
covered_loss <- modify_loss(
    loss = raw_loss, 
    deductible = DEDUCTIBLE, 
    limit = LIMIT, 
    coinsurance = COINSURANCE, 
    per.loss = FALSE 
)

p2 <- ggplot2::ggplot(
    data = data.frame(
        'loss_amt' = c(raw_loss, covered_loss), 
        'loss_type' = c(
            rep('Incurred Raw Loss', times = length(raw_loss)), 
            rep('Insurer Covered Loss', times = length(covered_loss))
        )
    ), 
    mapping = ggplot2::aes(x = loss_amt, fill = loss_type, color = loss_type)
) + 
    ggplot2::geom_histogram(ggplot2::aes(y=..density..), alpha = 0.5, position = 'identity') + 
    ggplot2::geom_density(alpha = 0.2) + 
    ggplot2::xlim(LOSS_MIN, LOSS_MAX) + 
    ggplot2::facet_wrap(~ loss_type, ncol = 1) + 
    ggplot2::labs(
        # axis labels 
        x = 'Loss Amount ($)', 
        y = 'Density'# , 
        # # legend title 
        # fill = 'Type', color = 'Type'
    ) + 
    ggplot2::theme_bw() + 
    # do not show legend 
    ggplot2::theme(legend.position = "none")

grid::grid.newpage()
grid::grid.draw(rbind(
    ggplot2::ggplotGrob(p1), 
    ggplot2::ggplotGrob(p2)
))
```

# Case Study 

## Policy Specification 

For this example, let's consider an insurance policy with 

- (Ordinary) Policy Deductible: \$ 5,000 
- Policy Limit: \$ 20,000 
- Coinsurance Factor: 90 \%

```{r policy_param}
# policy params 
DEDUCTIBLE <- 5000 
LIMIT <- 20000 
COINSURANCE <- 0.9 

# helper function 
coverage_modification <- function(x) modify_loss(
    loss = x, 
    deductible = DEDUCTIBLE, 
    limit = LIMIT, 
    coinsurance = COINSURANCE, 
    per.loss = FALSE 
)
```

## Data Simulation 

Suppose the incurred raw loss amount follows an unknown data generating process 
$$
L \sim \ln \mathcal{N}(\mu, \sigma)
$$
with the following parameters 
```{r data_params}
# lognormal distribution param 
MU <- 9 
SIGMA <- 1 
```

Let's simulate some data for this case study 
```{r data_sim}
# data simulation 
loss_raw <- rlnorm(n = n_sample, meanlog = MU, sdlog = SIGMA) 
loss_mod <- coverage_modification(loss_raw)
```

## Data Simulation (Cont.)

```{r plot, echo=FALSE, message=FALSE}
ggplot2::ggplot(
    data = data.frame(
        'loss_amt' = c(loss_raw, loss_mod), 
        'loss_type' = c(
            rep('Incurred Raw Loss', times = length(loss_raw)), 
            rep('Insurer Covered Loss', times = length(loss_mod))
        )
    ), 
    mapping = ggplot2::aes(x = loss_amt, fill = loss_type, color = loss_type)
) + 
    ggplot2::geom_histogram(ggplot2::aes(y=..density..), alpha = 0.5, position = 'identity') + 
    ggplot2::geom_density(alpha = 0.2) + 
    ggplot2::facet_wrap(~ loss_type, ncol = 1) + 
    ggplot2::labs(
        # axis labels 
        x = 'Loss Amount ($)', 
        y = 'Density'# , 
        # # legend title 
        # fill = 'Type', color = 'Type'
    ) + 
    ggplot2::scale_x_continuous(labels = scales::comma) + 
    ggplot2::scale_y_continuous(labels = scales::scientific) + 
    ggplot2::theme_bw() + 
    # do not show legend 
    ggplot2::theme(legend.position = "none")
```

## Parameter Estimation 

Affine transformation of modified loss data alone won't be sufficient. (Statistics omitted here.) Instead, we should transform the probability density function according to the contract parameters. 

For modified loss $y \in (0, \alpha (u-d))$, 
$$
F_Y(y) = \frac{
    F_X \big( \frac{y + \alpha d}{\alpha(1+r)} \big) 
    - F_X \big( \frac{d}{1+r} \big)
}{
    1 - F_X \big( \frac{d}{1+r} \big)
}
$$

```{r cdf}
# specify only CDF of loss_raw to get CDF of loss_mod
ploss_mod <- actuar::coverage(
    cdf = plnorm,  
    deductible = DEDUCTIBLE, 
    limit = LIMIT, 
    coinsurance = COINSURANCE, 
    franchise = FALSE, 
    inflation = 0, 
    per.loss = FALSE 
)
```

## Parameter Estimation (Cont.) 

For modified loss $y \in (0, \alpha (u-d))$, 
$$
f_Y(y) = 
\frac{1}{\alpha(1+r)} \cdot 
\frac{
    f_X \big( \frac{y + \alpha d}{\alpha(1+r)} \big) 
}{
    1 - F_X \big( \frac{d}{1+r} \big)
}
$$
with a point probability mass at $y = \alpha (u - d)$, 
$$
\frac{
    1 - F_X \big( \frac{u}{1+r} \big)
}{
    1 - F_X \big( \frac{d}{1+r} \big)
}
$$

```{r pdf}
# specify both PDF and CDF of loss_raw to get PDF of loss_mod
dloss_mod <- actuar::coverage(
    pdf = dlnorm, cdf = plnorm, 
    deductible = DEDUCTIBLE,
    limit = LIMIT, 
    coinsurance = COINSURANCE, 
    franchise = FALSE, 
    inflation = 0, 
    per.loss = FALSE 
)
```

## Parameter Estimation (Cont.) 

```{r start_ft, echo=FALSE}
start_fit <- fitdistrplus::fitdist(data = loss_mod, distr = 'lnorm')$estimate
start_est <- lapply(
    split(start_fit, names(start_fit)), 
    unname
)
```

By supplying our customized density function `dloss_mod` (which captures coverage modifications), we can estimate the parameters of raw loss using `fitdistrplus::fitdist` as before. 

\footnotesize
```{r}
# parameter fit  
loss_mod_fit <- fitdistrplus::fitdist(
    data = loss_mod, 
    distr = dloss_mod,
    start = start_est, 
    method = 'mle'
)
loss_mod_fit
```
\normalsize

## Concluding Remarks 

We can fit transformed data using existing approaches if we can construct the corresponding distribution functions. 

- [`actuar`](https://cran.r-project.org/web/packages/actuar/index.html): Actuarial Functions and Heavy Tailed Distributions 
  - `actuar::coverage`: Density and Cumulative Distribution Function for Modified Data 
- [`fitdistrplus`](https://cran.r-project.org/web/packages/fitdistrplus/index.html): Help to Fit of a Parametric Distribution to Non-Censored or Censored Data 
  - `fitdistrplus::fitdist`: Fit of Univariate Distributions to Non-Censored Data
