---
title: | 
  | AXIS/ORMIR Student Challenge
  |   Utility for Fitting Loss Distributions
short-title: "Distribution Fitting Utility"
author: "Haoen CUI and Peng JIN"
short-author: "H. CUI & P. JIN"
date: '`r format(Sys.Date(), "%B %d, %Y")`'      # Month DD, YYYY (Main Slide)
short-date: '`r format(Sys.Date(), "%m/%d/%Y")`' # MM/DD/YYYY (Lower Right)
institute: "University of Illinois at Urbana-Champaign"
short-institute: "UIUC"
department: "Department of Mathematics"          # Institute must be defined
section-titles: false                            # Provides slide headings
safe-columns: true                               # Enables special latex macros for columns.
header-includes:
   - \usepackage{bbm}                            # for indicator variable
output: 
   uiucthemes::beamer_illinois
---

# Agenda
### Agenda

\begincols

\column{0.40\linewidth}

**AXIS's Demand**

- Distribution Fitting
  + Coverage Modification 
  + Changing Volume 
  + Binned Data 
- Documentation 
  + Usable 
  + Maintainable 
  + Scalable

\column{0.08\linewidth}

$$\boldsymbol{\Longrightarrow}$$

\column{0.50\linewidth}

**Agenda** 

- Statistical Models 
  + Coverage Modifications of Severity Random Variables 
  + Coverage Modifications of Frequency Random Variables 
  + CDF Interpolation 
  + Modeling Loss Ratio via GLM
- Demonstration 
- Documentation 
- Conclusion 

\endcols

# Statistical Models
## Severity Random Variables
### Coverage Modifications of Severity Random Variables

\begincols

\column{0.59\linewidth}

- Definition 
  + $L$, raw loss size random variable
    - Likelihood: $f_L(l;\boldsymbol{\theta})$
  + $Y$, coverage modified loss size random variable
    - Likelihood: $f_Y(y;\boldsymbol{\theta})$
    - Dataset: $\mathcal{D}$

\column{0.40\linewidth}

- Coverage Modifications 
  + Deductible (ordinary/franchise) 
  + Limit 
  + Per loss / per payment 
  + Coinsurance factor $\alpha$ 
  + Interest rate $r$

\endcols

#### Data Fitting Procedure for Coverage Modified Severity Random Variables

\begin{align*}
\text{Parametric Model} \quad f_L(l;\boldsymbol{\theta}) &\xrightarrow{\text{modifications}} f_Y(y;\boldsymbol{\theta}) \xrightarrow{\text{generates}} \mathcal{D}=\{y_i\}_{i=1}^n \\
&\xrightarrow{\text{est methods}} \widehat{\boldsymbol{\theta}} \\
&\xrightarrow{\text{evaluate}} \text{GoF Stats/Plots, Boot Conf Interval} 
\end{align*}

## Frequency Random Variables
### Coverage Modifications of Frequency Random Variables

- Definition 
  + $N^L$, raw loss count random variable
  + $N^P$, **deductible** modified **payment** count random variable

\begin{exampleblock}{Theorem: $N^L$ and $N^P$ are from the same parametric family under conditions}
Assume compound frequency model, $N^P = \sum_{i=1}^{N^L} \mathbbm{1}\{L>d\}$ where $N^L$ is parametrized by $(\alpha, \beta)$ and $\nu := \mathbb{P}(L>d)$. Let $G(\cdot)$ denote the probability generating function of a random variable. 
Then, \textit{under some conditions}, 
$$G_{N^P}(z) = G_{N^L}(z;\alpha^{*}, \beta^{*})$$
where change of parameter is given by 
$$\alpha = \mathbb{P}(N^L = 0) \qquad \alpha^{*} = \mathbb{P}(N^P = 0) \qquad \beta^{*} = \nu \beta$$
\end{exampleblock}


### Coverage Modifications of Frequency Random Variables

| $N^L$ with parameters $\boldsymbol{\theta}=(\alpha, \beta)$    | $N^P$ with parameters $\boldsymbol{\theta}^{*}=(\alpha^{*}, \beta^{*})$   | 
| :------------------------------------ | :------------------------------------------- | 
| Poisson($\lambda$)     | $\lambda^{*}=\nu\lambda$          | 
| Zero-Modified Poisson($p^M_0$, $\lambda$)    |  ${p^M_0}^{*} = \frac{1-e^{-\lambda^{*}}}{1-e^{-\lambda}}p^M_0 + \frac{e^{-\lambda^{*}} - e^{-\lambda}}{1-e^{-\lambda}}$ and $\lambda^{*}=\nu\lambda$ | 

Table: Examples for Change of Parameters Under Coverage Modifications

#### Data Fitting Procedure for Coverage Modified Frequency Random Variables

\begin{align*}
& \text{Parametric Model } (a,b,0) \text{ or } (a,b,1) \quad p_{N^L}(n;\boldsymbol{\theta}) \text{ and } \nu = \mathbb{P}(L > d) \\ 
&\xrightarrow{\text{modifications}} p_{N^P}(n;\boldsymbol{\theta}^{*}) \xrightarrow{\text{generates}} \mathcal{D}=\{n_i\}_{i=1}^n \\
&\xrightarrow{\text{est methods}} \widehat{\boldsymbol{\theta}}^{*} \xrightarrow{\text{transform}} \widehat{\boldsymbol{\theta}} \xrightarrow{\text{evaluate}} \text{GoF Stats/Plots, Boot Conf Interval} 
\end{align*}



## CDF Interpolation
### Binned Data and CDF Interpolation

- Definition 
  + $L$, raw loss size random variable 
  + Dataset $\mathcal{D}=\{F_n(l_b)\}_{b=1}^B$ where $(l_b)_{b=1}^B$ is the set of pre-determined break points
- Goal 
  + Find $\hat{F} \in \mathcal{F}$ based on data $\mathcal{D}$ 
  + where $\mathcal{F}$ is constrained such that its members are all valid distributions 

#### Data Fitting Procedure for Global Parametric Family $\mathcal{F}_{\theta}$

Suppose $\mathcal{F}_{\theta}$ is a *global* (*not piecewise*) parametric family, then 
$$\hat{F}_{\theta} = \underset{F_{\theta} \in \mathcal{F}_{\theta}}{\operatorname{argmin}} d(F_{\theta}, F_n)$$
where $d(\cdot, \cdot)$ is a distance measure between distributions (e.g., Kullback–Leibler divergence) and $F_n$ is empirical distribution based on the observed dataset $\mathcal{D}$. 

### Binned Data and CDF Interpolation

In general, we may consider to interpolate the dataset $\mathcal{D}$ to get our estimate $\hat{F}$. However, we will need to ensure the constraint for arbitrary $\mathcal{F}$. There is no equivalent conditions for a general class of functions. 

- Polynomials: *The Bernstein Form of a Polynomial* by Cargo and Shisha (1966) provides a sufficient condition for an increasing quantile function modeling by a polynomial of degree $n$  
- Piecewise Pareto: Linear relationship between log(rank) and log(loss)
$$\forall l \geq l_0 \qquad \ln \bigg( \mathbb{P}(L > l) \bigg)  = - \alpha \ln (l) + \alpha \ln (l_0) $$
- Hunter and Drown (2017) implemented some methods to interpolate CDFs 
  + `R` package: [`binsmooth`](https://cran.r-project.org/web/packages/binsmooth/index.html) 
  + Publication: *Better Estimates from Binned Income Data: Interpolated CDFs and Mean-Matching*, available on [arXiv](https://arxiv.org/abs/1709.09705)

## Loss Ratio
### Modeling Loss Ratio via GLM

- Tweedie distribution $(\mu, \phi, p)$
  + Poisson($\lambda$) distributed sum of Gamma(shape$=\alpha$, scale$=\theta$) random losses
  + Tweedie mean, $\mu = \lambda \alpha \theta$ 
  + Tweedie poewr, $p = \frac{1}{\alpha + 1} + 1$
  + Tweedie dispersion, $\phi = \frac{\lambda^{1-p}(\alpha \theta)^{2-p}}{2-p}$
- CAS Monograph *Generalized Linear Models for Insurance Rating* suggests to model loss ratio (aka pure premium) using GLM with Tweedie distribution
  + Note that Tweedie GLM implicitly assumes frequency and severity moves in the same direction 

# Demonstration
### Coverage Modifications of Severity Random Variables

\begincols

\column{0.49\linewidth}

\begin{figure} 
\centerline{\includegraphics[width=0.95\textwidth]{img/PolicySetup.png}}
\caption{Policy Configuration}
\end{figure}

\column{0.49\linewidth}

\begin{figure} 
\centerline{\includegraphics[width=0.95\textwidth]{img/ModelPlot.png}}
\caption{Model Plot with Estimated Parameters}
\end{figure}

\endcols

### Adding User-Defined Distributions

- Run User-Defined Distribution in *`R` Code*
```{r eval = FALSE} 
# R default: Define Distribution
source(helper_file.R)
dbeta
pbeta
rbeta
# helper.R: Coverage Modification
actuar::coverage(pdf = ddist, cdf = pdist, 
                 deductible = deductible, limit = limit, 
                 franchise = franchise, inflation = inflation, 
                 coinsurance = coinsurance,
                 per.loss = per.loss)
# helper.R: Fit Distribution 
fitdistrplus::fitdist(data = temp_data_raw, distr = dist.name, 
                      method = method)
```

### Adding User-Defined Distributions

- Run User-Defined Distribution in *`R` Shiny*
```{r eval = FALSE}
# helper.R: Add to the list of distributions 
Distribution.list <- list(
  "select.beta"=c(ddist = dbeta, pdist = pbeta, 
                  rdist = rbeta, qdist = qbeta, 
                  dist.name = "beta", dist.nparams = 2), 
  ...)
# App new.R: Add to User Interface
box(title = "Fitting Specification",
    selectInput(inputId = "loss.dist", 
                "Select distribution of raw loss random variable(L)",
                c("Beta"="select.beta", ...))
```

# Conclusion

## Documentation
### Summary of Documentations

- *User Manual*: Step-by-step procedure to run the utility 
- *Statistics Manual*: Details of statistical assumptions, results, and procedures 
- *Maintainability* 
  + Users can add new loss distributions 
  + Code formatted and commented under Google `R` style 
  + Reproducibility ensured by `R` project and `packrat`
- *Scalability* 
  + Modular and functional programming 
  + Ability to add user-defined loss distributions 


### Model Constrains and Directions of Future Works

- Credibility framework (e.g., Bayesian)
- From parametric to non-parametric 
- Cross Validation 

### Conclusion 

\begincols

\column{0.35\linewidth}

**AXIS's Demand**

- Distribution Fitting
  + Coverage Modification 
  + Changing Volume 
  + Binned Data 
- Documentation 
  + Usable 
  + Maintainable 
  + Scalable

\column{0.05\linewidth}

$$\boldsymbol{\Longrightarrow}$$

\column{0.58\linewidth}

**Project Deliverables** 

- Statistical Models 
  + Coverage Modifications 
    - Severity Random Variables 
    - Frequency Random Variables 
  + CDF Interpolation 
  + Modeling Loss Ratio via GLM
- Platform
  + Interface: `R Shiny`
  + Development: `R` project and `packrat`
- Documentation 
  + User Manual 
  + Statistics Manual 


\endcols

# Reference 
### References

- G. E. Willmot, H. H. Panjer, S. A. Klugman. (2012). *Loss Models: From Data to Decisions, 4th Edition*.  
- M. B. Finan. (2017). *An Introductory Guide in the Construction of Actuarial Models: A Preparation for the Actuarial Exam C/4*. Retrieved From [Link](http://faculty.atu.edu/mfinan/actuarieshall/CGUIDE.pdf).
- D. Bahnemann. (2015). *Distributions for Actuaries*. Retrieved From  [Link](http://www.casact.org/pubs/monographs/index.cfm?fa=bahnemann-monograph02)  
- M. Goldburd, A. Khare, and D. Tevet. (2016). *Generalized Linear Models for Insurance Rating*. Retrieved From  [Link](https://www.casact.org/pubs/monographs/papers/05-Goldburd-Khare-Tevet.pdf)  
